Implicit yaddayadda

### 2011-10-27 papers/PLDI2012/implicitization.tex
On ne distingue plus X_i^s et X_i^d
Le fait d'avoir deux valeurs abstraites pour chaque point est utile pour
le narrowing, et le fait de moralement couper les points en deux est
importante dans les formules du SMT-solving (b^s et b^d).

Mais pour la découverte de nouveaux chemins, on (Julien et moi) ne trouve
pas de cas où c'est intéressant de mettre à jour une valeur temporaire
X^d plutôt que de mettre à jour directement X^s (en faisant une mise à
jour sans élargissement).

Le code fait en réalité la distinction, donc pas exactement ce que fait
l'algo après ce commit, mais la distinction X^s/

Guided static analysis \cite{DBLP:conf/sas/GopanR07} applies to the transition
graph of the program. We now present a new technique applying this analysis on the implicit
multigraph from \cite{Monniaux_Gonnord_SAS11}, thus avoiding control flow merges with
unfeasible paths.
In this section, we use the same notations as \ref{sec:path_focusing}, except
that we call the abstract values $X_i^s$ instead of $X_i$ to leave room for
another abstract value $X_i^d$ later in the algorithm.

The combination of these two techniques aims at first discovering a precise
inductive invariant for a subset of paths between two points in $P_R$, 
by the mean of ascending and narrowing iterations. When an
inductive invariant has been found, we add new feasible paths to the subset and
compute an inductive invariant for this new subset, starting with the results
from the previous analysis.
In other words, our technique considers an ascending sequence of
subsets of the paths between two points in $P_R$.
We iterate the operations until we consider the whole program, i.e all the
feasible paths. The result will then be an inductive invariant of the entire program.

@@

We distinguish three phases in the main loop of the analysis:
\begin{enumerate}
\item \label{step:addingpaths} We start finding a new relevant subset
  $P \cup P'$ of the graph.
  Either the previous iteration or the initialization lead us to a
  state where there are no more paths in the previous subset $P$,
  starting at $p_i$, that make the abstract values of the successors
  grow (otherwise, the SMT solver would not have answered
  ``\emph{unsat}''). Narrowing iterations preserve this property.
  However, there may exist such paths in the entire multigraph, that
  are not in $P$. This phase computes these paths and adds them to
  $P'$ (line~\ref{alg=computeNewPaths} in
  algorithm~\ref{algo:combined}). We also use SMT solving to discover
  these new paths, but we slightly change the SMT formula given to the
  SMT solver: we now simply check for the satisfiability of $f(p_i)$
  instead of $f(p_i) \wedge P$. Since we already know that $f(p_i)
  \wedge P$ is unsatisfiable, all the paths given by the SMT solver
  won't be in $P$. This phase is described in~\ref{subsec:addingpaths}
  and corresponds to lines in~\ref{alg=start-add-paths} to
  \ref{alg=end-add-paths} algorithm~\ref{algo:combined}.
\item \label{step:ascending} Given a new subset $P$, we search for paths starting at point
  $p_i \in P_R$, such that these paths are in $P$, i.e are included in
  the working subgraph. Each time we find a path, we update the
  abstract value of the destination point of the path. This is the
  phase explained in~\ref{subsec:ascending}, and corresponds to
  lines~\ref{alg=start-ascending} to~\ref{alg=end-ascending} in
  algorithm~\ref{algo:combined}.
\item  \label{step:narrowing} We perform narrowing iterations the usual way
  (line~\ref{alg=narrowing} in algorithm~\ref{algo:combined}) and
  reiterate from step 1 unless there are no more points to explore,
  i.e. $A' = \emptyset$.
\end{enumerate}

@@

@@

@@

In order to find which path to focus on, we construct an SMT formula $f(p_i)$, whose
model when satisfiable is a path that starts in $p_i$, goes to a successor $p_j
\in P_R$ of $p_i$, such that the image of $X_{i}^s$ by the path transformation
is not included in the current $X_{j}^s$.
Intuitively, such a path makes the abstract value $X_{j}^s$ grow, and thus is
an interesting path to focus on. We loop until the formula becomes unsatisfiable,
meaning that the analysis of $p_i$ is finished.

If we note $Succ(i)$ the set of indices $j$ such that $p_j \in P_R$ is a
successor of $p_i$ in the expanded multigraph, and $X_i^s$ the abstract value
associated to $p_i$ :
$$f(p_i) = \rho \wedge b_i^s \wedge 
\displaystyle\bigwedge_{j \in P_R \atop j \neq i} \neg
b_j^s \wedge x \in X_i^s \wedge \displaystyle\bigvee_{j \in Succ(i)} (b_j^d \wedge
x \notin X_j^s)$$

@@

Our technique computes the fixpoint iterations on a ascending sequence of
subgraphs, until the complete graph is reached.

When the analysis of a subgraph is finished, meaning that the abstract values
for each control points has converged to an inductive invariant for this subgraph,
the next subgraph to work on has to be computed.

This new subgraph is the union of the paths of the previous one with a set
$P'$ of new paths that become feasible regarding the current abstract values.

\MM{Cette partie est devenue toute petite, il y avait un résumé global
  de l'algo ici. Ça me parait suffisant comme ça}

It is not necessary to add into $P'$ all the new feasible paths. We just have
to add them one after another, until no more path can make the invariant grow,
as proposed by Algorithm \ref{algo:computepaths}. For that, we use another
abstract value associated to a control point $p_j$, noted $X_j^d$, which is
distinct from $X_j^s$, and initialized to $X_j^s$.

@@

\begin{figure}[h]
  \begin{center}
    \input{./techniques.tex}
  \end{center} 
  \vspace{-20px}
  \caption{Comparison of the abstract values obtained by classical abstract
  interpretation (S), \emph{Guided Static
  Analysis} (G), \emph{Path-focused} technique (PF), our combined technique
  (G+PF), and its version with disjunctive invariants (DIS).}
  \label{fig:techniques}
\end {figure}

>>>>

Guided static analysis \cite{DBLP:conf/sas/GopanR07} applies to the transition
graph of the program. We now present a new technique applying this analysis on the implicit
multigraph from \cite{Monniaux_Gonnord_SAS11}, thus avoiding control flow merges with
unfeasible paths.
In this section, we use the same notations as \ref{sec:path_focusing}.

The combination of these two techniques aims at first discovering a precise
inductive invariant for a subset of paths between two points in $P_R$, 
by the mean of ascending and narrowing iterations. When an
inductive invariant has been found, we add new feasible paths to the subset and
compute an inductive invariant for this new subset, starting with the results
from the previous analysis.
In other words, our technique considers an ascending sequence of
subsets of the paths between two points in $P_R$.
We iterate the operations until the whole program (i.e all the
feasible paths) has been considered. The result will then be an
inductive invariant of the entire program.

@@

We distinguish three phases in the main loop of the analysis:
\begin{enumerate}
\item \label{step:addingpaths} We start finding a new relevant subset
  $P \cup P'$ of the graph.
  Either the previous iteration or the initialization lead us to a
  state where there are no more paths in the previous subset $P$,
  starting at $p_i$, that make the abstract values of the successors
  grow (otherwise, the SMT solver would not have answered
  ``\emph{unsat}''). Narrowing iterations preserve this property.
  However, there may exist such paths in the entire multigraph, that
  are not in $P$. This phase computes these paths and adds them to
  $P'$. This phase is described in~\ref{subsec:addingpaths}
  and corresponds to lines in~\ref{alg=start-add-paths} to
  \ref{alg=end-add-paths} algorithm~\ref{algo:combined}.
\item \label{step:ascending} Given a new subset $P$, we search for paths starting at point
  $p_i \in P_R$, such that these paths are in $P$, i.e are included in
  the working subgraph. Each time we find a path, we update the
  abstract value of the destination point of the path. This is the
  phase explained in~\ref{subsec:ascending}, and corresponds to
  lines~\ref{alg=start-ascending} to~\ref{alg=end-ascending} in
  algorithm~\ref{algo:combined}.
\item  \label{step:narrowing} We perform narrowing iterations the usual way
  (line~\ref{alg=narrowing} in algorithm~\ref{algo:combined}) and
  reiterate from step 1 unless there are no more points to explore,
  i.e. $A' = \emptyset$.
\end{enumerate}

@@

\MM{Pas sûr que ça soit nécessaire, mais vu qu'on a eu une version
  fausse avant, c'est peut-être bien de dire comment on est arrivés à
  la meilleure version.}
The order of steps is important: narrowing has to be performed before
adding new paths, or some spurious new paths would be added to $P$,
and starting with the addition of new paths avoids wasting time
doing the ascending iterations on an empty graph.

@@

\MM{Maintenant que l'ordre a changé dans l'algo, on peut peut-être
  échanger cette section et la suivante ?}

@@

In order to find which path to focus on, we construct an SMT formula $f(p_i)$, whose
model when satisfiable is a path that starts in $p_i$, goes to a successor $p_j
\in P_R$ of $p_i$, such that the image of $X_{i}$ by the path transformation
is not included in the current $X_{j}$.
Intuitively, such a path makes the abstract value $X_{j}$ grow, and thus is
an interesting path to focus on. We loop until the formula becomes unsatisfiable,
meaning that the analysis of $p_i$ is finished.

If we note $Succ(i)$ the set of indices $j$ such that $p_j \in P_R$ is a
successor of $p_i$ in the expanded multigraph, and $X_i$ the abstract value
associated to $p_i$ :
$$f(p_i) = \rho \wedge b_i^s \wedge 
\displaystyle\bigwedge_{j \in P_R \atop j \neq i} \neg
b_j^s \wedge x \in X_i \wedge \displaystyle\bigvee_{j \in Succ(i)} (b_j^d \wedge
x \notin X_j)$$

@@

Our technique computes the fixpoint iterations on a ascending sequence of
subgraphs, until the complete graph is reached.
When the analysis of a subgraph is finished, meaning that the abstract values
for each control points has converged to an inductive invariant for this subgraph,
the next subgraph to work on has to be computed.

This new subgraph is the union of the paths of the previous one with a set
$P'$ of new paths that become feasible regarding the current abstract values.
The paths in $P'$ are computed one after another, until no more path
can make the invariant grow. This is line~\ref{alg=computeNewPaths} in
Algorithm~\ref{algo:combined}, which corresponds to
Algorithm~\ref{algo:computepaths}. We also use SMT solving to discover
these new paths, but we subtly change the SMT formula given to the
SMT solver: we now simply check for the satisfiability of $f(p_i)$
instead of $f(p_i) \wedge P$. Since we already know that $f(p_i)
\wedge P$ is unsatisfiable, none of the paths given by the SMT solver
will be in $P$.

@@

\begin{figure}[h]
  \MM{Y a-t-il moyen d'avoir un truc plus lisible en impression noir
    et blanc, genre des hachures ?}
  \begin{center}
    \input{./techniques.tex}
  \end{center} 
  \vspace{-20px}
  \caption{Comparison of the abstract values obtained by classical abstract
  interpretation (S), \emph{Guided Static
  Analysis} (G), \emph{Path-focused} technique (PF), our combined technique
  (G+PF), and its version with disjunctive invariants (DIS).}
  \label{fig:techniques}
\end {figure}
